Metagenomic Barcoding of Pollen Loads Offers Insights on the Foraging Patterns of Queen Bumble Bees

### 2023-02-19 script/manuscript_MEE.Rmd
added new graphs, reporting on phenology, now working on "expert" recoding of results =-\

To estimate the duration of dates in which plant species were flowering weibull estimates of several phenological parameters all spatially modelled taxa were developed (@belitz2020accuracy, @pearse2017statistical). Only BIEN records which occurred in the Omernik Level 4 Ecoregions within 15km of the study area (n = 5 Level 4 Ecoregions, or conditionally 6 ecoregions if enough records were not be found in the nearest 5), and which were from herbarium records were included. To remove temporally irrelevant herbarium records, i.e. material collected during times which flowering is impossible at the study area due to snow cover, we used the SnowUS data set (@iler2021conceptual, @tran2019cloud) from 2000-2017 were analyzed for the first three days of contiguous snow absence, and the first three days of contiguous snow cover in Fall. Herbarium records after the 3^rd^ quantile for melt, and the 1^st^ quantile for snow cover of these metrics were removed. Species with > 10 records had their weibull distributions generated for the date when 10% of individuals had begun flowering, when 50% were flowering, and when 90% of individuals had flowered, we used the initiation and cessation dates, respectively, as effective start and ends of flowering.

@@

ggplot(data=correct, aes(x=Observation, y=value,
                         group = Bombus.Species, fill = Bombus.Species)) +
 # geom_line(linewidth = 1.25, linetype = 'dotted', alpha = 0.3)+
  geom_jitter(size = 3, shape = 23, aes(color = Bombus.Species), 
              height = 0.05, width = 0.3) +
  scale_fill_manual('Bee species', values = bee_cols) +
  scale_color_manual('Bee species', values = bee_cols) +
  theme_bw() +
  scale_y_continuous(breaks = c(0,2,4,6,8,10), limits = c(0,10))+
  labs(title = 'Taxa shared between the top 10 observations and molecular work', 
       y = 'Shared Observations', x= NULL) + 
  theme(panel.grid.major.x = element_blank(),
        panel.border = element_blank())

@@

Only 58 of these 388 species (n = 34.568, Mdn = 31) were able to be compared to plot based observational data from the long term (1974–2012) data set (@caradonna2014shifts). Of these species relatively high accord was observed between the long-term ground truthed data set, and the modelled species. There was very strong evidence that the weibull estimates were positively associated with the observed onset (r^2^ = 0.72, p < 0.0001, tau = 0.61) and peak (r^2^ = 0.70, p < 0.0001, tau = 0.65) of flowering, and that the number of herbarium samples had a moderate effect on the estimates (p = 0.004 and p = 0.034 respectively). There was very strong evidence that the weibull estimates had a positive association with the observed cessation of flowering (r^2^ =  0.4339,  p < 0.0001, tau = 0.489), however their was no evidence that sample size had an effect (p = 0.349). There was moderate evidence that the weibull estimates, with an effect of sample size, had a weak positive association with the observed duration of flowering (p = 0.0401, r^2^ = 0.07, tau = 0.17).

```{r plot phenology estimates, echo=F, fig.cap="Modelled dates of when major flowering events occurred", out.width = '100%'}
knitr::include_graphics("../graphics/plots/phenology_estimates_10_90.png")
# these are the tenth and nintey  percent non-parametric confidence intervals. 
```

@@

... with samples 22, and 41 both having < 100 reads indicates, virtual failure of these records (REMOVE from analyses).

@@

The relationship between the number of pollen grains in a sample and the number of sequence reads is roughly *curvilinear*, where grains which are present in trace amounts are overestimated by sequence counts, while grains present in high amounts are underestimated. This is likely due to the proportion of high false positives which occur in the classification process with next-generation sequencing (@bell2021comparing). There was evidence of a strong correlation between the proportion of grains per morphotype and the number of sequences per group (`r round(ct[["estimate"]][["tau"]], 3)`, p < 0.0001, n = `r nrow(pol_p)`).

@@

e <- blst %>% 
  filter(Family %in% c('Boraginaceae', 'Hydrophyllaceae', 'Ranunculaceae', 
                       'Celastraceae', 'Fabaceae')) %>% 
  select(sample.id, Genus) %>% 
  mutate(Present = 1) %>% 
  group_by(Genus) %>% 
  distinct() %>% 
  pivot_wider(names_from = Genus, values_from = Present) %>% 
  mutate(across(.col = everything(), ~replace_na(.,  0))) %>% 
  column_to_rownames('sample.id')

@@

To ascertain the extent to which records of multiple species in a family, which were suspected to be sampling artefacts occurred in molecular samples an index of similarity, ala jaccard, the affinity index was used to assess co-occurrence (@mainali2022better, @mainali2022alpha). Numerous taxa from the family Ranunculaceae Jussieu (*Caltha* L. sp., *Thalictrum* L. spp., *Trollius* L. sp., *Aquilegia* L. spp.), had $\alpha$ scores which indicated that they are only present when a more common confamilial taxa *Delphinium barbeyi* (Huth) Huth *nuttallianum* Pritz. were recorded. A similar relationship was observed in the Hydrophyllaceae R.Br. with samples placed in *Nemophila* Nutt., which only occurred when the more abundant *Hydrophyllum* L. species were present.

@@

```{r}

blst_final <- blst %>% 
  mutate(TAXON_EXPERTS = case_when(
      TAXON_NEW %in% c('Cynoglossum amplifolium', 'Cynoglossum anchusoides', 
      'Cynoglossum officinale') ~ 'Mertensia',
      TAXON_NEW %in% c('Caltha leptosepala', 'Trollius europaeus', 'Trollius') ~ 'Delphinium',
    # 'Thalictrum thalictroides' ~ 'Delphinium',
      TAXON_NEW == 'Erigeron' ~ 'Erigeron grandiflorus', 
      TAXON_NEW == 'Scabrethia scabra' ~ 'Wyethia arizonica',
      TAXON_NEW == 'Nemophila breviflora' ~ 'Hydrophyllum',
      TAXON_NEW == 'Erigeron grandiflorus' ~ 'Erigeron grandiflorus',
      TAXON_NEW == 'Epilobium canum' ~ 'Chamaenerion angustifolia',
      TAXON_NEW == 'Paxistima myrsinites' ~ 'Parnassia palustris', # original was probably right !!! shows the consequence of bad time estimates.#, 
    # TRUE ~ as.character(TAXON_NEW)
  )) %>% 
  select(TAXON, TAXON_NEW, TAXON_EXPERTS)

# no TARAXACUM OFICINALE, shows a limitation of the data sources.

Situations where SDM's led to incorrect results at the species level are evident with classification to *Scabrethia Scabra* (Hooker) W.A. Weber, this match almost certainly representing *Wyethia arizonica* A. Gray (@weber1998new), a taxon known to be visited by Queen bee's via our floral observations.

It is not unlikely that much of the difference in the results between the observational and molecular work are attributable to the challenges in detecting rare events in these smaller sizes. For example, no more than 10 bee corbiculae loads per species were sequenced with the Mdn = `r median(n_corbiculae)` ... , and the median of interactions with the top 5 plant sizes constituted `r median(prop_obs)`  of the top.

An expected inaccuracy of the classification scheme is in genus level placements, e.g. were *Epilobium* L. (Onagraceae Juss.) spp. were classified. However, given the small size of their flowers in the study area, these results more likely indicate that a species of *Chamaenerion* Seg. (a segregate genus) such as *C. angustifolium* (L.) Scop.  or *latifolium* (L.) Sweet is occasionally utilized, as it supported by limited palynology data.

Accordingly, combining the results of floral observations, and palynology, molecular sequencing - both pre and post processing, we subjectively developed re-classifications of the contents of pollen grains...

@@

##### ~ What we DEMONSTRATED ~

We have demonstrated how the Angiosperms533 hyb-seq probes may be used for plant barcoding in a metagenomic context (@johnson2019universal, @hollingsworth2016telling). This was exemplified in an ecologically relevant scenario, where the results have immediate implications for natural history driven fundamental science and land management. The test pollen loads contained a number of closely related taxa, some in notoriously morphologically difficult clades with rapid rates of diversification (e.g. *Mertensia*, *Lupinus* L.), at naturally occurring proportions (@nevado2016widespread, @nazaire2014phylogenetic).  We incorporated spatial and temporal approaches for creating custom sequence databases an approach which is readily applicable to any lab group with the capacity to perform next-generation sequencing across the entirety of multiple continents, and which we expect to be highly beneficial in many study areas. By combining insights from these novel approaches with an extensive observational field based study we show how these methods may be applied to test a variety of hypotheses related to ecological interactions.

##### ~ What CHALLENGES we FACED ~

  The SDM's which we generated, with relatively few occurrence records and few modelling iterations, performed beyond expectations, likely due to the utility of the predictor variables and strong alignment of vegetation by orographic precipitation in the study area. However, we had difficulties in evaluating our predictions in an operational context. We utilized the database query approach, to only model species with a high probability of not being dispersal limited to the focal area, and focused on a relevant subset of many of these species ranges to reduce the contributions of range wide adaptions on habitat (@sork2018genomic, @joshi2001local). While the models worked well compared to both test, and validation with external point data, moving from points to polygon features was more difficult. We were able to compare our results to 1) a Flora, 2) lists of plants used by Bumble Bees at plots; the former inappropriate in that it contained a great number of species which we sought to use modelling to reduce *e.g.* all strictly alpine species, and the latter inappropriate in that it contained only species relevant to *Bombus* but had no official 'absence' data. Further given the, size of the minimum spanning tree (AREA???) which we extracted points to, a formal floristic inventory would still be a time intensive process. Accordingly, we expect the real results of our data lay somewhere in between these two evaluations; with an excess of species predicted present (@dubuis2011predicting, @calabrese2014stacking, @pinto2021predicting), but few enough that they lend themselves to metabarcoding. We observe that our models seemed very capable of effectively identifying alpine species and removing them in binomial contexts. 
  Difficulties in temporal models related to variability in drivers of flowering phenology.

##### ~ WHAT we learned about Bee foraging (BASIC) ~

@@

Results from palynological analyses show that several species of bee show near perfect fidelity to the genus *Hydrophyllum* on a per visit basis... General results show high congruence between foraging and molecular results, indicating that concerns regarding mismatch between observational networks need not persit with *Bombus* studies...

##### ~ WHAT this tells us about Bee foraging (APPLIED) ~

@@

##### ~ WHERE we see spatial/temporal going

  We have concerns regarding the number of persons training to become and practice botany, and grave concerns regarding the funding mechanisms for floristic and field based botanical research and for centralized authorities to produce consensus opinions on alpha taxonomy (@prather2004decline, @kramer2015report, @prather2004implications, @crisci2020end, @manzano2021flippant, @stroud2022botanical). To reduce the effects of a low population density of botanists on the maintenance of and production of flora's and to foster meta-genomics across landscapes without field stations we utilized Species Distribution Modelling to generate predictive species lists. In this proof of concept example we performed several iterations of modelling runs, and several approaches (i.e. the 'linear models', and the 'machine learning'), which took notable amounts of compute power. We suspect the possible deleterious nature of this endeavor may be reduced by: 1) more field surveying by crews will reduce the need to generate as many species 2) fewer runs of models, 3) only running machine learning models which do not require an explicitly process to reduce spatial autocorrelation. However, given the time required to perform all aspects of a study, even our amount of computation was negligible. Further, we are very optimistic about the possibility for persons to perform these tasks, as mentioned we utilized roughly only one quarter of the records which were digitally available for presence, and we suspect others will have enough records to perform this process nearly anywhere else in the temperate. In certain scenarios modelling of predicted species via more formally tailored S(tacked)-SDM or J(oint)-SDM approaches may be beneficial (@wilkinson2021defining, @pinto2021predicting, @schmitt2017ssdm).

@@

##### ~ WHERE we see MOLECULAR going

>>>>

To estimate the duration of dates in which plant species were flowering weibull estimates of several phenological parameters all spatially modelled taxa were developed (@belitz2020accuracy, @pearse2017statistical). Only BIEN records which occurred in the Omernik Level 4 Ecoregions within 15km of the study area (n = 5 Level 4 Ecoregions, or conditionally 6 ecoregions if enough records were not be found in the nearest 5), and which were from herbarium records were included. To remove temporally irrelevant herbarium records, i.e. material collected during times which flowering is impossible at the study area due to snow cover, we used the SnowUS data set (@iler2021conceptual, @tran2019cloud) from 2000-2017 were analyzed for the first three days of contiguous snow absence, and the first three days of contiguous snow cover in Fall. Herbarium records after the 3^rd^ quantile for melt, and the 1^st^ quantile for snow cover of these metrics were removed. Species with > 10 records had their weibull distributions generated for the date when 10% of individuals had begun flowering, when 50% were flowering, and when 90% of individuals had flowered, we used the initiation and cessation dates, respectively, as effective start and ends of flowering. These estimates were compared to groundtruthed data from a long term experiment 1974-2012, and the field data from 2015, using kendall's tau.

@@

ggplot(data=correct, aes(x=Observation, y=value,
                         group = Bombus.Species, fill = Bombus.Species)) +
  geom_jitter(size = 3, shape = 23, aes(color = Bombus.Species), 
              height = 0.05, width = 0.3) +
  scale_fill_manual('Bee species', values = bee_cols) +
  scale_color_manual('Bee species', values = bee_cols) +
  theme_bw() +
  scale_y_continuous(breaks = c(0,2,4,6,8,10), limits = c(0,10))+
  labs(title = 'Taxa shared between the top 10 observations and molecular work', 
       y = 'Shared Observations', x= NULL) + 
  theme(panel.grid.major.x = element_blank(),
        panel.border = element_blank())

@@

Only 58 of these 388 species (n = 34.6, Mdn = 31) were able to be compared to plot based observational data from the long term (1974–2012) data set (@caradonna2014shifts). Of these species relatively high accord was observed between the long-term ground truthed data set, and the modelled species. There was very strong evidence that the weibull estimates were positively associated with the observed onset (p < 0.0001, tau = 0.61),  peak (p < 0.0001, tau = 0.65), and cessation of flowering (p < 0.0001, tau = 0.49). There was moderate evidence that the weibull estimates had a weak positive association with the observed duration of flowering (p = 0.58, tau = 0.17).

Of the previous 58 species compared, 47 of these could be compared to plot based data from the six sites observed in 2015. Due to methodological differences, the peak flowering was not compared, and due to the low performance of attempts to model 'duration' in the previous step it was also not compared. There was very strong evidence that the weibull estimates were positively associated with the observed onset (p < 0.0001, tau = 0.58),  and cessation of flowering (p < 0.0001, tau = 0.40).

```{r plot phenology estimates, echo=F, fig.cap="Modelled dates of when major flowering events occurred compared between long term and modelled data", out.width = '100%'}
knitr::include_graphics("../graphics/plots/phenology_estimates_10_90.png")
# these are the tenth and nintey  percent non-parametric confidence intervals. 
```

```{r plot phenology estimates JEO, echo=F, fig.cap="Modelled dates of when major flowering events occurred compared between 2015 and modelled data", out.width = '100%'}
knitr::include_graphics("../graphics/plots/phenology_estimates-jeo_10_90.png")
```

@@

@@

The relationship between the number of pollen grains in a sample and the number of sequence reads is roughly linear, where grains which are present in trace amounts are overestimated by sequence counts, while grains present in high amounts are underestimated. This is likely due to the proportion of high false positives which occur in the classification process with next-generation sequencing (@bell2021comparing). There was evidence of a strong correlation between the proportion of grains per morphotype and the number of sequences per group (`r round(ct[["estimate"]][["tau"]], 3)`, p < 0.0001, n = `r nrow(pol_p)`).

@@

e <- blst %>% 
  filter(Family %in% c('Boraginaceae', 'Hydrophyllaceae', 'Ranunculaceae', 
                       'Celastraceae')) %>% 
  select(sample.id, Genus) %>% 
  mutate(Present = 1) %>% 
  group_by(Genus) %>% 
  distinct() %>% 
  pivot_wider(names_from = Genus, values_from = Present) %>% 
  mutate(across(.col = everything(), ~replace_na(.,  0))) %>% 
  column_to_rownames('sample.id')

@@

To ascertain the extent to which records of multiple species in a family, which were suspected to be sampling artefacts occurred in molecular samples an index of similarity, ala jaccard, the affinity index was used to assess co-occurrence (@mainali2022better, @mainali2022alpha). Numerous taxa from the family Ranunculaceae Jussieu (*Caltha* L. sp., *Thalictrum* L. spp., *Trollius* L. sp., *Aquilegia* L. spp.), had $\alpha$ scores which indicated that they are only present when a more common confamilial taxa *Delphinium barbeyi* (Huth) Huth *nuttallianum* Pritz. were recorded. A similar relationship was observed in the Hydrophyllaceae R.Br. with samples placed in *Nemophila* Nutt., which only occurred when the more abundant *Hydrophyllum* L. species were present. The size of flower of *Nemophila breviflora* A. Gray make it unlikely to be visited by Bumble Bees, and it is a false positive. The floral morphology of *Thalictrum*  spp. also makes it unlikely to be visited (BELEIVE JANE SAID THIS- IF NOT STRIKE OUT), and while evidence of visits to Caltha and Trollius are lacking, due to the association between the reads these results appear unlikely.

@@

```{r Build reference dataset for reclassification process}

pollen_sample_info <- readxl::read_xlsx(file.path('../data', 
  list.files(path = '../data', pattern = 'xlsx') )) %>% 
  select(site, date, sample.id) %>% 
  distinct() %>% 
  filter(!date %in% c('42170', '42169'))

files <- list.files(path = '../data')

flr_date_ranges <- read.csv(
  file.path('../data',  files[grep('flower_ranks', files)] )) %>% 
  select(site, doy, week, plant.species:abun.rank) %>% 
  group_by(plant.species, site) %>% 
  select(site, week, doy, plant.species) %>% 
  mutate(Genus = str_extract(plant.species, '^.+?(?=[.]|$)')) %>% 
  mutate(Genus = case_when(
    Genus == 'Erythrocoma' ~ 'Geum', 
    Genus == 'Amerosedum' ~ 'Sedum', 
    Genus == 'Pentaphylloides' ~ 'Dasiphora',
    TRUE ~ Genus
    )) %>% 
  mutate(plant.species = case_when(
    plant.species == 'Lathyrus.leucanthus' ~ 'Lathyrus.lanszwertii',
    plant.species == 'Lupinus.bakeri' ~ 'Lupinus.sericeus',
    TRUE ~ plant.species
  )) %>% distinct()
 # this provides us with a data.frame of the ACTUAL flowering dates!

week_tab <- flr_date_ranges %>% 
  ungroup() %>% 
  select(doy, week) %>% 
  distinct()

flr_date_ranges <- select(flr_date_ranges, -doy)

blst_expert <- blst %>% 
  select(species, sample.id, TAXON_NEW, Genus, Family, Prcnt_seqs, Prcnt_Seqs_Class) %>% 
  mutate(RecordID = 1:n()) %>% 
  left_join(., pollen_sample_info,  by = 'sample.id') %>% 
  relocate(site:date, .before = TAXON_NEW) %>% 
  mutate(doy = lubridate::yday(date)) %>% 
  left_join(., week_tab, by = 'doy') %>% 
  left_join(., flr_date_ranges, by = c('site', 'Genus', 'week')) %>% 
  relocate(sample.id, .before = plant.species)

  # reduce to temporally similar results

ranunc_splash <- c('Caltha', 'Trollius', 'Thalictrum', 'Aquilegia')

# all samples with duplicate genera in them # be careful these are not easy mappings
duplicate_reps_in_sample <- blst_expert %>%

  # identify the duplicate samples
  drop_na(plant.species) %>% 
  group_by(sample.id, TAXON_NEW) %>% 
  ungroup(TAXON_NEW)  %>% 
  distinct(plant.species, .keep_all = T) %>% 
  group_by(sample.id, TAXON_NEW) %>% 
  filter(n() >= 2) %>%

  # now split the reads assigned to them into each constituent taxon ~ 'unwise King Soloman'
  ungroup() %>% 
  group_by(RecordID) %>% 
  mutate(Prcnt_seqs_Reclass = Prcnt_seqs/n())

blst_reclass_expert <- blst_expert %>% 
  filter(!RecordID %in% duplicate_reps_in_sample$RecordID) %>% 
  rowwise() %>% 
  mutate(Genus = str_trim(Genus))  %>%

  mutate(TAXON_EXPERT = case_when(
    sample.id == 3 & Genus == 'Delphinium' ~ plant.species,
    sample.id == 4 ~ plant.species, # 4 has a viola split
    sample.id == 4 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 6 ~ plant.species, # has a delphinium split
    sample.id == 6 & Genus %in% ranunc_splash ~ 'Delphinium.barbeyi',

    sample.id == 9 ~ plant.species, # has both Delphinium
    sample.id == 10 ~ plant.species,

    sample.id == 14 ~ plant.species,

    sample.id == 18 ~ plant.species,
    sample.id == 19 ~ plant.species,
    sample.id == 20 ~ plant.species,
    sample.id == 21 ~ plant.species,
    sample.id == 23 ~ plant.species, # has both Viola
    sample.id == 23 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 24 ~ plant.species,
    sample.id == 24 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 25 ~ plant.species,
    sample.id == 25 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 26 ~ plant.species,
    sample.id == 26 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
  #  sample.id == 27 ~ plant.species, # has both Hydrophyllum
    sample.id == 27 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 28 ~ plant.species,
    sample.id == 28 & Genus %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 29 ~ plant.species,
    sample.id == 30 ~ plant.species,
    sample.id == 30 & Genus == 'Cynoglossum' ~ 'Mertensia.fusiformis', 
    sample.id == 32 ~ plant.species,
    sample.id == 32 & TAXON_NEW %in% ranunc_splash ~ 'Delphinium.nuttallianum',
    sample.id == 33 ~ plant.species, # has both Viola

    sample.id == 35 ~ plant.species, # has both Viola
    sample.id == 35 & TAXON_NEW %in% ranunc_splash ~ 'Delphinium.nuttallianum',

    sample.id == 36 ~ plant.species,

    sample.id == 39 ~ plant.species, # has both hydrophyllum
    sample.id == 39 & TAXON_NEW %in% ranunc_splash ~ 'Delphinium.nuttallianum',

    sample.id == 41 ~ plant.species,
    sample.id == 41 & Genus == 'Cynoglossum' ~ 'Mertensia.ciliata',

    sample.id == 42 ~ plant.species,
    sample.id == 43 ~ plant.species,
    #sample.id == 44 ~ plant.species, # 44 has both delphinium

    sample.id == 46 ~ plant.species,
    sample.id == 49 ~ plant.species,
    sample.id == 49 & TAXON_NEW %in% ranunc_splash ~ 'Delphinium.barbeyi',
  #  sample.id == 51 ~ plant.species, # 51 has a viola split
    sample.id == 51 & TAXON_NEW %in% ranunc_splash ~ 'Delphinium.nuttallianum',

    sample.id == 52 ~ plant.species,
    sample.id == 54 ~ plant.species,
    sample.id == 55 ~ plant.species,

    sample.id == 56 ~ plant.species,
    sample.id == 56 & Genus == 'Nemophila' ~ 'Hydrophyllum fendleri',

    sample.id == 60 ~ plant.species,
    sample.id == 65 ~ plant.species, # 65 has a mertensia split

   TRUE ~ TAXON_NEW
  )) %>%

  # use this to sum up the total sequences by original read record
  distinct(sample.id, TAXON_NEW, .keep_all = T) %>% 
  group_by(sample.id, plant.species) %>% 
  mutate(Prcnt_seqs_Reclass = sum(Prcnt_seqs)) %>%

  # now reduce to the amount of rows of the proper classification
  ungroup() %>% 
  distinct(sample.id, TAXON_EXPERT, .keep_all = T)

  mutate(TAXON_EXPERTS = case_when(
      TAXON_NEW %in% c('Cynoglossum amplifolium', 'Cynoglossum anchusoides', 
      'Cynoglossum officinale', 'Cynoglossum pringlei') ~ 'Mertensia',
      TAXON_NEW %in% c('Caltha leptosepala', 'Caltha palustris', 'Trollius europaeus', 'Trollius', 
                       'Thalictrum thalictroides', 'Aquilegia ecalcarata', 
                       'Thalictrum tuberosum') ~ 'Delphinium',
      TAXON_NEW == 'Erigeron' ~ 'Erigeron grandiflorus', 
      TAXON_NEW == 'Scabrethia scabra' ~ 'Wyethia arizonica',
      TAXON_NEW == 'Nemophila breviflora' ~ 'Hydrophyllum',
      TAXON_NEW == 'Epilobium canum' ~ 'Chamaenerion angustifolia',
      TAXON_NEW == 'Paxistima myrsinites' ~ 'Parnassia palustris', #, 
      TAXON_NEW == 'Taraxacum cucullatum' ~ 'Taraxaum oficinale'
    # original was probably right !!! shows the consequence of bad time estimates.#, 
  #   TRUE ~ as.character(TAXON_NEW)
  )) %>% 
  select(TAXON, TAXON_NEW, TAXON_EXPERTS)

rm(pollen_sample_info)
```

Situations where SDM's led to incorrect results at the species level are evident with classification to *Scabrethia Scabra* (Hooker) W.A. Weber, this match almost certainly representing *Wyethia arizonica* A. Gray (@weber1998new), a taxon known to be visited by Queen bee's via our floral observations. An expected inaccuracy of the classification scheme is in genus level placements, e.g. were *Epilobium* L. (Onagraceae Juss.) spp. were classified. However, given the small size of their flowers in the study area, these results more likely indicate that a species of *Chamaenerion* Seg. (a segregate genus) such as *C. angustifolium* (L.) Scop.  or *latifolium* (L.) Sweet is occasionally utilized, as it supported by limited palynology data. An issue with reclassification within the family level in combination with time included reclassifying *Parnassia palustris* L. to *Paxistima myrsinites* (Pursh) Rafinesque. However, based on flower size it is more likely that the visited taxon was *P. palustris*.

It is not unlikely that much of the difference in the results between the observational and molecular work are attributable to the challenges in detecting rare events in these smaller sizes. For example, no more than 10 bee corbiculae loads per species were sequenced with the Mdn = `r median(n_corbiculae)` ... , and the median of interactions with the top 5 plant sizes constituted `r median(prop_obs)`  of the top interactions. 
Accordingly, combining the results of floral observations, and palynology, molecular sequencing - both pre and post processing, we subjectively developed re-classifications of the contents of pollen grains...

@@

We have demonstrated how the Angiosperms533 hyb-seq probes may be used for plant barcoding in a metagenomic context (@johnson2019universal, @hollingsworth2016telling). This was exemplified in an ecologically relevant scenario, where the results have immediate implications for natural history guided fundamental science and land management. The test pollen loads contained a number of closely related taxa, some in notoriously morphologically difficult clades with rapid rates of diversification (e.g. *Mertensia*, *Lupinus* L.), at naturally occurring proportions (@nevado2016widespread, @nazaire2014phylogenetic). We incorporated spatial and temporal approaches for creating custom sequence databases an approach which is readily applicable to any lab group with the capacity to perform next-generation sequencing across the entirety of multiple continents, and which we expect to be highly beneficial in many study areas. By combining insights from these novel approaches with an extensive observational field based study we show how these methods may be applied to test a variety of hypotheses related to ecological interactions.

  The SDM's which we generated, with relatively few occurrence records and few modelling iterations, performed beyond expectations, likely due to the utility of the predictor variables and strong alignment of vegetation by orographic precipitation in the study area. However, we had difficulties in evaluating our predictions in an operational context. We utilized the database query approach, to only model species with a high probability of not being dispersal limited to the focal area, and focused on a relevant subset of many of these species ranges to reduce the contributions of range wide adaptions on habitat (@sork2018genomic, @joshi2001local). While the models worked well compared to both test, and validation with external point data, moving from points to polygon features was more difficult. We were able to compare our results to 1) a Flora, 2) lists of plants used by Bumble Bees at plots; the former inappropriate in that it contained a great number of species which we sought to use modelling to reduce *e.g.* all strictly alpine species, and the latter inappropriate in that it contained only species relevant to *Bombus* but had no official 'absence' data. Further given the, size of the minimum spanning tree  which we extracted points to, a formal floristic inventory would still be a time intensive process. Accordingly, we expect the real results of our data lay somewhere in between these two evaluations; with an excess of species predicted present (@dubuis2011predicting, @calabrese2014stacking, @pinto2021predicting), but few enough that they lend themselves to metabarcoding. We observe that our models seemed very capable of effectively identifying alpine species and removing them in binomial contexts. 
  Difficulties in temporal models related to variability in drivers of flowering phenology. AND SMALL SAMPLE SIZE

@@

Results from palynological analyses show that several species of bee show near perfect fidelity to the genus *Mertensia* on a per visit basis... General results show high congruence between foraging and molecular results, indicating that concerns regarding mismatch between observational networks need not persit with *Bombus* studies...

@@

  We have concerns regarding the number of persons training to become and practice botany, and grave concerns regarding the funding mechanisms for floristic and field based botanical research and for centralized authorities to produce consensus opinions on alpha taxonomy (@prather2004decline, @kramer2015report, @prather2004implications, @crisci2020end, @manzano2021flippant, @stroud2022botanical). To reduce the effects of a low population density of botanists on the maintenance of and production of Flora's and to foster meta-genomics across landscapes without field stations we utilized Species Distribution Modelling to generate predictive species lists. In this proof of concept example we performed several iterations of modelling runs, and several approaches (i.e. the 'linear models', and the 'machine learning'), which took notable amounts of compute power. We suspect the possible deleterious nature of this endeavor may be reduced by: 1) more field surveying by crews will reduce the need to generate as many species 2) fewer runs of models, 3) only running machine learning models which do not require an explicitly process to reduce spatial autocorrelation. However, given the time required to perform all aspects of a study, even our amount of computation was negligible. Further, we are very optimistic about the possibility for persons to perform these tasks, as mentioned we utilized roughly only one quarter of the records which were digitally available for presence, and we suspect others will have enough records to perform this process nearly anywhere else in the temperate. In certain scenarios modelling of predicted species via more formally tailored S(tacked)-SDM or J(oint)-SDM approaches may be beneficial (@wilkinson2021defining, @pinto2021predicting, @schmitt2017ssdm).

@@
